### 2014-04-02 paper/designspace.tex

\section{The Design Space of Pipes}
By using different openings, designs, and topologies of pipes, and inserting different media into the completed pipes, inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

@@

Pipes can have four opening styles: open, return, semi-closed, and fully enclosed (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access, e.g., via soundwaves.

@@

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create electronic circuitry.  An open pipe filled with conductive paint can power an LED (see Figure \ref{fig:openings}A).

@@

Semi-closed pipes are open on the system side, closed on the user side.  The closed interface may be fabricated from a malleable material, e.g., a rubber membrane, and attached to a fluid pump to create haptic feedback (see Figure \ref{fig:openings}C).

@@

\subsection{Design of Pipes}
Pipe design may emphasize either exterior connection points or interior paths.  Focused on the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive. Alternatively, the neon sign's output in Figure \ref{fig:neon} is based on the \emph{path} of the pipe.

@@

Pipe network topologies enable different interactions.  Splitting or mixing pipes offer flexibility in output, allowing a single source to have multiple outputs or multiple sources to combine (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out.

\subsection{Media in Pipes}

Pipes can be filled with a variety of media to create different interface affordances and capabilities.

@@

Incompressible fluids (``liquids'') perform similarly to gases.  Additionally, using driable conductive liquids like copper paint, pipes can become custom wires.

>>>>

\section{The Design Space of Pipes}
\tovi{clean up writing}By using different openings, designs, and topologies of pipes, and inserting different media into the completed pipes, inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

@@

Pipes can have four opening styles: open, return, semi-closed, and fully enclosed (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access, e.g., via soundwaves. \tovi{still not convinced we need this distinction}

@@

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create electronic circuitry.  For example, an open pipe filled with conductive paint can power an LED (see Figure \ref{fig:openings}A).

@@

Semi-closed pipes are open on the system side, closed on the user side. A potential use of this is to create tactile output: The closed interface may be fabricated from a malleable material, e.g., a rubber membrane, and the open end attached to a fluid pump to create haptic feedback (see Figure \ref{fig:openings}C).

@@

\subsection{Design of Pipes}
Pipe design may emphasize either exterior connection points or interior paths.  Focused on the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive. Alternatively, the neon sign's output in Figure \ref{fig:neon} is based on the \emph{path} of the pipe. \tovi{still think we can remove this dimension}

@@

Pipe network topologies enable different interactions.  Splitting or mixing pipes offer flexibility in output, allowing a single source to have multiple outputs or multiple sources to combine (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out. \tovi{give simple examples of where this might be useful}

\subsection{Media in Pipes}
\tovi{clean up this writing}
One of the most interesting aspects of the design space is that pipes can be filled with a variety of media to create different interface affordances and capabilities. We consider gas, liquids, solids, particulates, and threadables.

@@

Incompressible fluids (``liquids'') perform similarly to gases.  Additionally, using driable conductive liquids like copper paint, pipes can become custom wires. \tovi{custom water jet devices (uist 2013 competition)}

### 2014-04-02 paper/exampleobjects.tex

To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using the tool.  All prototypes were fabricated in a single piece, unless described otherwise.

@@

We created a set of touch-sensitive toys and a companion app, reminiscent of the boat application in Acoustic Barcodes \cite{Harrison-acoustic}.  The brain, rabbit, and boat each have several distinct touch points on their surfaces.  These touch points are connected by an interior star topology of tubes filled with conductive copper paint, and touch sensing is performed via a single wire and SFCS (see Figure \ref{fig:toys}).  We built a smart base which can distinguish between the toys and also determine which toy is mounted: since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.

The models we used for the toys' exteriors were all downloaded from the free website Thingiverse.  All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running code by Instructables users DZL and madshobye.

\subsection{Breathing Bunny (semi-closed, gas, exterior)}

We created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:breathe}).  For this, we used a combination air/vacuum pump: one terminal creates a vacuum while the other creates positive pressure.  Our rabbit has two tubes, one open tube exiting at its mouth and one semi-closed tube capped with rubberlike material in its abdomen.  We connected one tube to each of our pump's terminals, and using a programmable power supply we can mimic a rabbit's breathing pattern.  This example was printed on our Objet in rubberlike material, with support flushed post-print. \valkyrie{this paragraph is really unclear}

@@

We created a maze game with a single particle trapped in a fully enclosed tube.  This maze is based on an SVG we created by hand and processed with our internal path tube tool.  To fabricate the tube, we created it in two halves that were fastened together via glue, due to limitations in our ability to remove support material otherwise.  We created this prototype on our Makerbot in two halves, inserting a marble before closing.

>>>>

To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using a Series of Tubes.  All prototypes were fabricated in a single piece, unless described otherwise.

@@

We created a set of touch-sensitive toys and a companion app, reminiscent of the boat application in Acoustic Barcodes \cite{Harrison-acoustic} \bjoern{People will not  know what this is; they shouldn't have to look up that paper to understand. I rewrote:}. The objects --- a brain, rabbit and boat, in our example --- can be set on a base, which announces the type of object, and also the names of special features of an object when that feature is touched. For example, touching a strip on top of the brain model yields the announcement ``primary motor cortex''. The distinct touch points on each object are connected by an interior star topology of tubes filled with conductive copper paint, and touch sensing is performed via a single wire and SFCS (see Figure \ref{fig:toys}).  We built a smart base which can distinguish between the toys and also determine which toy is mounted: since each toy and each gesture has a distinct capacitive signature, we use a simple classifier trained to detect both toy and gesture based on profile.

The models we used for the toys' exteriors were all downloaded from the free website Thingiverse.  All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running code by Instructables users DZL and madshobye \bjoern{what is that code? Maybe just mention open source SFCS code and footnote a URL?}.

\subsection{Breathing Bunny (semi-closed, gas, exterior)}
\bjoern{Start each section with the larger point why we built each example. `` (semi-closed, gas, exterior)'' is likely to short. ``To demonstrate how gases can be used to both provide sensation to the user through openings and to deform a model internally, we created a rabbit...''}
\bjoern{why do we have two bunnies in this and the prior application? Maybe we should substitute a different model for the bunny above.}

We created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:breathe}) \bjoern{How does it simulate breathing? Be specific ``the bunny exhales and its abdominal area rises and falls with each breath'' or something like that}.  For this, we used a combination air/vacuum pump: one terminal creates a vacuum while the other creates positive pressure.  Our rabbit has two tubes, one open tube exiting at its mouth and one semi-closed tube capped with rubberlike material in its abdomen.  We connected one tube to each of our pump's terminals, and using a programmable power supply we can mimic a rabbit's breathing pattern.  This example was printed on our Objet in rubberlike material, with support flushed post-print. \valkyrie{this paragraph is really unclear}

@@

We created a maze game with a single particle trapped in a fully enclosed tube.  This maze is based on an SVG we created by hand and processed with our internal path tube tool.  To fabricate the tube, we created it in two halves that were fastened together via glue, due to limitations in our ability to remove support material otherwise. \bjoern{observant reviewers will call us out on this - that many of our techniques may not actually be workable in a single piece because you can't get the support material out. I think if you had a lye bath you could actually get more complicated models done, as long as you're patient and there's some entry and exit for the lye to reach all pipes.}  We created this prototype on our Makerbot in two halves, inserting a marble before closing.

### 2014-04-02 paper/limitations.tex

Some media are difficult to insert in pipes, or may not be compatible with all types of pipes.  Fluids in particular provide challenges.  For example, the conductive paint we used in our designs can take multiple days to dry in tubes of sufficient length.  Any liquid kept in tubes printed by the Objet can cause the exterior of the part to discolor due to seepage.  Hobbyist machines do not guarantee watertightness of their prints, so leaking of water or air can occur.

>>>>

Some media are difficult to insert in pipes, or may not be compatible with all types of pipes.  Fluids in particular provide challenges. \bjoern{I'd think that solids would be harder - they can get stuck along the way.} For example, the conductive paint we used in our designs can take multiple days to dry in tubes of sufficient length.  Any liquid kept in tubes printed by the Objet can cause the exterior of the part to discolor due to seepage.  Hobbyist machines do not guarantee watertightness of their prints, so leaking of water or air can occur.

\bjoern{also comment on the difficulty of getting support materials out?}
